---
title: "Module 2"
output: html_document
date: '2022-10-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ISLR)
library(dplyr)
library(leaps)
library(car)
library(devtools)
```

## Body Fat Data

```{r}
bodyfat <- read.csv('BodyFat.csv')
```

## Step 1. Analyzing Raw Data

```{r}
head(bodyfat)
tail(bodyfat)
```

```{r}
summary(bodyfat)
```

```{r}
hist(bodyfat$AGE)
hist(bodyfat$WEIGHT)
bodyfat %>% subset(bodyfat$WEIGHT > 250)

hist(bodyfat$HEIGHT)
bodyfat %>% subset(bodyfat$HEIGHT < 50)

hist(bodyfat$BODYFAT)
hist(bodyfat$NECK)
bodyfat %>% subset(bodyfat$NECK > 50)

hist(bodyfat$CHEST)
bodyfat %>% subset(bodyfat$CHEST > 130)

hist(bodyfat$ABDOMEN)
bodyfat %>% subset(bodyfat$ABDOMEN > 140)

hist(bodyfat$HIP)
bodyfat %>% subset(bodyfat$HIP > 120) # may have to get rid of another outlier row 41

hist(bodyfat$THIGH)
bodyfat %>% subset(bodyfat$THIGH > 75)

hist(bodyfat$KNEE)
bodyfat %>% subset(bodyfat$KNEE > 4) # 244th observation

hist(bodyfat$ANKLE)
bodyfat %>% subset(bodyfat$ANKLE<= 30) # 31, 86

hist(bodyfat$BICEPS)
bodyfat %>% subset(bodyfat$BICEPS > 40)

hist(bodyfat$FOREARM)
bodyfat %>% subset(bodyfat$FOREARM < 25) # no obv outliers

hist(bodyfat$WRIST)
bodyfat %>% subset(bodyfat$WRIST > 21) # 39, 41

# most all have row 39 which is deleted bc weight is huge outlier

max(bodyfat$BODYFAT)
min(bodyfat$BODYFAT)
sort(bodyfat$BODYFAT)

ggplot(bodyfat) + geom_histogram(aes(ABDOMEN), binwidth = 5, color = 'black', fill = 'blue') + 
    labs(title ="Histogram of Abdomen Circumference", 
             x = "Abdomen Circumference (cm)", y = "Count") + 
  theme(plot.title = element_text(hjust = 0.5))
```



### Clean data set
```{r}
bodyfat_out <- bodyfat %>% subset(bodyfat$BODYFAT >= 4 & bodyfat$BODYFAT <= 40 & bodyfat$WEIGHT < 300 & bodyfat$HEIGHT > 50 & bodyfat$ANKLE <= 30 & bodyfat$WRIST <= 21) %>% select(-ADIPOSITY, -IDNO, -DENSITY) %>% mutate(HEIGHT = HEIGHT * 2.54, WEIGHT = WEIGHT*2.2)
```

```{r}
cor(bodyfat_out)
bodyfat_out <- bodyfat_out %>% mutate(ABDOMEN_scale = (1/6) * ABDOMEN)
```

## Modeling

UPDATED: fitting fill model first with bodyfat_out data
```{r}
lm_out <- lm(BODYFAT ~., data = bodyfat_out)
summary(lm_out)
plot(lm_out, which=4, cook.levels=cutoff)
plot(lm_out)
```

Best subsets:
```{r}
models <- regsubsets(BODYFAT ~ ., bodyfat_out, nvmax = 3)
models.sum <- summary(models)
models.sum

# Create four plots within a 2x2 frame to compare the different criteria
par(mfrow = c(2,2))
 # SSE
plot(models.sum$rss, xlab = "Number of predictors", ylab = "SSE", type = "l")
 # R2
plot(models.sum$adjr2, xlab = "Number of predictors", ylab = "Adjusted RSq", type = "l")
 # Mallow's Cp
plot(models.sum$cp, xlab = "Number of predictors", ylab = "Cp", type = "l")
 # BIC
plot(models.sum$bic, xlab = "Number of predictors", ylab = "BIC", type = "l")
```

Refit with top 3 features from best subsets:
```{r}
lm3 <- lm(BODYFAT ~ WEIGHT + ABDOMEN + WRIST, bodyfat_out)
summary(lm3)
vif(lm3)
```


Diagnostic plots for 3-factor model:
```{r}
# Save predicted values
bodyfat_out$predicted <- predict(lm3)
# Save residuals
bodyfat_out$resids <- residuals(lm3)
# Residuals vs Fitted values plot
ggplot(bodyfat_out, aes(x = predicted, y = resids)) + 
        geom_point() + geom_hline(yintercept = 0, color = "blue") + 
        labs(title ="Residuals versus Fitted values", 
             x = "Fitted values", y = "Residuals")
```

Residuals vs. abdomen predictor:
```{r}
# Residuals vs. Biomass predictor plot
ggplot(bodyfat_out, aes(x = ABDOMEN, y = resids)) + geom_point() + 
        geom_hline(yintercept = 0, color = "blue") + 
        labs(title ="Residuals versus Predictor of Biomass", x = "Biomass", 
             y = "Residuals")
```

Q-Q:
```{r}
# Normal probability plot
ggplot(bodyfat_out, aes(sample = resids)) + stat_qq() + stat_qq_line() +
 labs(title ="Normal probability plot",
 x = "Theoretical percentiles", y = "Sample percentiles") 
```


Some reduced models:
```{r}
lm2 <- lm(BODYFAT ~ WRIST + ABDOMEN, bodyfat_out)
summary(lm2)
vif(lm2)
```

```{r}
lm2 <- lm(BODYFAT ~ ABDOMEN + WEIGHT, bodyfat_out)
summary(lm2)
vif(lm2)
```

```{r}
lm_wrist <- lm(BODYFAT ~ WRIST, bodyfat_out)
summary(lm_wrist)
```

```{r}
lm_weight <- lm(BODYFAT ~ WEIGHT, bodyfat_out)
summary(lm_weight)
```


## FINAL MODEL:
```{r}
lm_ab <- lm(BODYFAT ~ ABDOMEN, bodyfat_out)
summary(lm_ab)
plot(lm_ab)
```

## Diagnostic Plots 
```{r}

eq <- function(df, x, y){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(R)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}


ggplot(bodyfat_out, aes(x=ABDOMEN, y=BODYFAT)) +
  geom_point() +   
  geom_smooth(method='lm', se=FALSE) + 
  geom_text(x = 80, y = 30, label = eq(bodyfat_out, bodyfat_out$ABDOMEN, bodyfat_out$BODYFAT ), parse = TRUE) +
  labs(title ="Regression Plot", 
             x = "Abdomen Circumference (cm)", y = "Bodyfat %") + 
  theme(plot.title = element_text(hjust = 0.5))
  
# code taken from https://www.roelpeters.be/how-to-add-a-regression-equation-and-r-squared-in-ggplot2/ for adding text into scatterplot
```

Residuals vs fitted values:
```{r}
# Save predicted values
bodyfat_out$predicted <- predict(lm_ab)
# Save residuals
bodyfat_out$resids <- residuals(lm_ab)
# Residuals vs Fitted values plot
ggplot(bodyfat_out, aes(x = predicted, y = resids)) + 
        geom_point() + geom_hline(yintercept = 0, color = "blue") + 
        labs(title ="Residuals versus Fitted values", 
             x = "Predicted % Bodyfat", y = "Residual Values") + 
  theme(plot.title = element_text(hjust = 0.5))
```

Residuals vs. abdomen predictor:
```{r}
# Residuals vs. predictor plot
ggplot(bodyfat_out, aes(x = ABDOMEN, y = resids)) + geom_point() + 
        geom_hline(yintercept = 0, color = "blue") + 
        labs(title ="Residuals versus Predictor of Abdomen Circumference", 
             x = "Abdomen Circumference (cm)", 
             y = "Residual Values") + 
  theme(plot.title = element_text(hjust = 0.5))
```


Q-Q:
```{r}
# Normal probability plot
ggplot(bodyfat_out, aes(sample = resids)) + stat_qq() + stat_qq_line() +
 labs(title ="Normal probability plot",
 x = "Theoretical percentiles", y = "Sample percentiles") +
  theme(plot.title = element_text(hjust = 0.5))
```



Forward stepwise? fit empty model first
```{r}
# Fit an empty model with only the response
FitStart <- lm(BODYFAT ~ 1, bodyfat_out)


# Fit a full model with all predictors
FitAll <- lm(BODYFAT ~ ., bodyfat_out)

# Run the stepwise regression with forward selection based on the AIC criterion
step(FitStart,direction="forward", scope = formula(FitAll))
```

** AIC is difficult to interpret- how we want to interpret this number, negative value being better
** using Rsquared as secondary metric to AIC??
** issues with collinearity; VIF threshold- 15?
** negative coefficients?
